Tomographic galaxy clustering with the Subaru Hyper Suprime-Cam first year public data release

### 2019-10-16 doc/Paper/main.tex
Updated plots and table with latest results.

\abstract{We analyze the clustering of galaxies in the first public data release of the Hyper Suprime-Cam Subaru Strategic Program. Despite the relatively small footprints of the observed fields, the data are an excellent proxy for the very deep photometric datasets such that the Large Synoptic Survey Telescope will acquire, and are therefore an ideal test bed for the analysis methods being implemented by the Dark Energy Science Collaboration. We select a magnitude limited sample with $i<24.5$ and analyze it in four tomographic redshift bins covering the range $0.15\lesssim z\lesssim1.5$. We carry out a Fourier-space analysis of the two-point clustering of this sample, including all auto- and cross-correlations between bins. We demonstrate the use of map-level deprojection methods to account for non-physical fluctuations in the galaxy number density caused by observational systematics. Through a halo occupation distribution analysis, we place constraints on the characteristic halo masses characterizing this sample as a function of redshift, finding a good fit up to scales $k_{\rm max}=1\,{\rm Mpc}^{-1}$ including both auto- and cross-correlations. Our results show monotonically decreasing average halo masses, which can be interpreted in terms of the drop-out of red galaxies at high redshifts for a flux-limited sample, reproducing previous results. In terms of photometric redshift systematics, we show that additional care is needed in order to marginalize over uncertainties in the redshift distribution in galaxy clustering, even for samples of this small size, and that these uncertainties can be significantly constrained by including cross-bin correlations. We are able to make a $\sim3\sigma$ detection of the effects of lensing magnification in the HSC data. When freeing up cosmological parameters, we find constraints on the amplitude of density fluctuations $\sigma_8$ and the cold dark matter abundance $\Omega_c$ that agree well with measurements from Planck and low-redshift probes. Finally, we use our analysis pipeline to study the clustering of galaxies as a function of limiting flux, and provide a simple model for the linear galaxy bias for magnitude limited samples as a function of limiting magnitude and redshift.}

@@

The past two decades have seen a revolution in our understanding of Universe and its constituents. While observations of the cosmic microwave background play a pivotal role \cite{Planck:2018}, they are only indirectly sensitive to physics of the low redshift universe, in particular expansion history and growth that are uniquely sensitive to properties of dark energy and modified gravity \cite{1903.12016}.

In studies of the low redshift universe, surveys of the large scale structure play an increasingly important role. Over a decade ago, the Dark Energy Task Force has divided the evolution of optical experiments into approximate stages \cite{0609591}. The current Stage III experiments including spectroscopic ones such as eBOSS \cite{1707.09322,1801.03038,1712.08064,1801.02689,1801.02891} and VIPERS \cite{1611.07048,1612.05645,1708.00026} and photometric ones such as DES \cite{1708.01531,1708.01530,1708.01536}, Hyper-SupremeCam \cite{1704.05858,2019PASJ...71...43H,1906.06041} and Kilo-Degree Survey / VIKING-450 \cite{1902.11265,1812.06077,Joudaki:2019} are approaching completion and the field is preparing for the Stage IV surveys, such as DESI \cite{1611.00036,1907.10688} and LSST \cite{0912.0201,1809.01669} and major satellite missions like Euclid \cite{1606.00180} and WFIRST \cite{1904.01174}.

Together with transformational sensitivity increases in the Stage IV surveys, the challenges of understanding and controlling systematic errors are becoming considerably more difficult \cite{1808.07335}. For photometric surveys, the deep surveys sensitive to many more sources inevitably come with crowded fields where effects of blending affect a significant fraction of all sources in the field. The increased blending leads to new uncertainties in isolating sources, measuring their fluxes (required for photometric redshifts) and inferring their shear estimates. These effects in turn lead to subtle sample selection effects, which are amplified by the interaction of the point spread function with blending \cite{1708.01533,1905.01324,1907.10572}. But even with perfect measurements of fluxes, our ability to infer redshift distribution of sources is hampered by incompleteness in spectroscopic samples used to calibrate redshifts \cite{1903.09325}. Finally, at the depths of Stage IV photometric surveys, instrumental and observational effects, such as the scattered light from bright objects, errors in the galaxy-star separation, reddedning by the milky way dust and others will impriment their own fluctuations into the galaxy density of objects. These fluctuations will, if not correctly accounted for, masquerade as intrinsic large-scale fluctuations in the cosmic density fields.  Therefore, it is likely that the Stage IV photometric experiments are likely to be limited by our ability to control systematics rather than the intrinsic statistical power of the survey.

Community is well aware of these challenges lying ahead. There are two main approaches that are being used in preparation for the arrival of data. On one hand, we are building very sophisticated realistic mock data sets, such as LSST DESC Data Challenge 2 \cite{1909.07340,1907.06530}. On the other hand, we are re-analyzing the existing precursor data to test-drive our methods and codes in realistic environment. In this paper we focus on the latter approach using Hyper Supreme-Cam Subaru Strategic Program (HSC-SSP) first data release  \cite{2018PASJ...70S...8A}. The main advantage of this dataset is that it is in many respects very similar to the LSST data, both in terms of depth and bands as well as primary data reduction and catalog generation. It covers essentially the same bands as LSST, with the exception of missing the UV $u$-band fluxes, which are unlikely to play a major role in cosmological data analysis. We will work with limiting magnitude for 5-$\sigma$ source detection of $i<24.5$, similar to expected magnitude cut to be used for LSST of $i<25.1$ after 1 year \cite{1809.01669}.  The primary data reduction codes used in HSC-SSP and planned for LSST are of the same lineage employing much of the same methods \cite{2018PASJ...70S...5B,1812.03248}. With the exception of a rather small sky area of around 90 square degrees used in this paper, the HSC-SSP data is a perfect example of how the LSST data will look like.

@@

The paper is structured as follows.  In Section \ref{sec:data} we discuss the HSC-SSP data and how we process it in order to generate a set of galaxy over-density maps together with maps of potential systematics. In Section \ref{sec:methods} we go through the methodology employed in our analysis of the data and in Section \ref{sec:results} we present results together with numerous tests. We discuss and conclude in Section \ref{sec:discussion}.

@@

  HSC-SSP is a photometric galaxy survey that has been awared 300 nights on the Subaru Telescope starting in 2014. The DR1 includes data from 61.5 nights covering and comes with fields of three depths: Wide (108 square degrees to $i\sim 26.4$), Deep (26 square degreesto $i\sim26.5$) and UltraDeep (4 square degrees to $i\sim 27$). In this work we focus on the Wide field, which has been observed in five broadband filters ($grizy$) and is distributed among 7 fields between 5 and 20 square degrees each. The image quality is impressive at median seeing of around 0.6 arcsec, which is considerably better than other comparable surveys (e.g. DES with around 0.9 arc sec) and also likely better than median seeing that will be achieved by  LSST. The data are processed with \texttt{hscPipe} \cite{2018PASJ...70S...5B} and  available to the community as a publicly available database.

@@

    \begin{table*}
      \caption{Summary of parameters varied in the MCMC with their respective priors. The posterior means and best-fits denote the values derived for our fiducial analysis. The uncertainties denote the $68 \%$ c.l..} \label{tab:params}
      \begin{center}
        \begin{tabular}{cccc}
          \hline\hline 
          Parameter & Prior & Posterior mean & Best-fit \\ \hline \Tstrut                             
          $M_{\mathrm{min}}$ & flat $\in [0., \,17.]$ & $11.89\substack{+0.24 \\ -0.25}$ & $11.76$ \\ 
          $M_{\mathrm{min}, p}$ & flat $\in [-10., \,10.]$ & $-0.0606\substack{+2.50 \\ -2.40}$ & $-2.95$ \\
          $M_{0}$ & flat $\in [0., \,15.]$ & $5.85\substack{+4.03 \\ -4.05}$ & $5.58$ \\
          $M_{0, p}$ & flat $\in [-5., \,10.]$ & $2.33\substack{+5.12 \\ -4.93}$ & $-0.925$ \\
          $M_{1}$ & flat $\in [0., \,15.]$ & $13.09\substack{+0.24 \\ -0.25}$ & $12.96$ \\ 
          $M_{1, p}$ & flat $\in [-10., \,15.]$ & $1.49\substack{+3.21 \\ -3.08}$ & $-2.07$ \\
          $\Delta z_{0}$ & flat $\in [-0.2, \,0.2]$ & $0.000147\substack{+0.102 \\ -0.096}$ & $-0.0456$ \\
          $\Delta z_{1}$ & flat $\in [-0.2, \,0.2]$ & $-0.0128\substack{+0.103 \\ -0.096}$ & $-0.0923$ \\
          $\Delta z_{2}$ & flat $\in [-0.2, \,0.2]$ & $-0.00170\substack{+0.109 \\ -0.104}$ & $-0.103$ \\
          $\Delta z_{3}$ & flat $\in [-0.2, \,0.2]$ & $0.0148\substack{+0.109 \\ -0.106}$ & $-0.0727$ \\
          $z_{w, 0}$ & flat $\in [-0.2, \,0.2]$ & $-0.0300\substack{+0.130 \\ -0.119}$ & $-0.177$ \\
          $z_{w, 1}$ & flat $\in [-0.2, \,0.2]$ & $0.000837\substack{+0.0780 \\ -0.0785}$ & $-0.0991$ \\
          $z_{w, 2}$ & flat $\in [-0.2, \,0.2]$ & $0.0342\substack{+0.0825 \\ -0.0794}$ & $0.0243$ \\
          $z_{w, 3}$ & flat $\in [-0.2, \,0.2]$ & $0.0355\substack{+0.122 \\ -0.138}$ & $0.151$ \\ \\
          $A_{\mu}$ & flat $\in [-3., \,3.]$ & $ - $ & $ - $ \\
          $\Omega_{c}$ & flat $\in [0.1, \,0.9]$ & $ - $ & $ - $ \\
          $\sigma_{8}$ & flat $\in [0.4, \,1.5]$ & $ - $ & $ - $ \\
          \hline\hline 
        \end{tabular}
      \end{center}
    \end{table*}

@@

Fig.~\ref{fig:constraints-fid-hod} shows the constraints on the HOD parameters $M_{\mathrm{min}}, \, M_{\mathrm{min}, p}, \, M_{1}, \, M_{1, p}$ for our fiducial model described in Sec.~\ref{ssec:methods.constr}. The constraints on all fitted parameters are shown in Fig.~\ref{fig:constraints-fid-full} and the corresponding best-fit values and means are shown alongside their $68 \%$ confidence limits in Tab.~\ref{tab:params}. As can be seen from Figures \ref{fig:constraints-fid-hod} and \ref{fig:constraints-fid-full}, the data allow us to constrain $M_{\mathrm{min}}(z)$ and $M_{1}(z)$ whereas $M_{0}(z)$ is unconstrained\footnote{In the following, we therefore only show constraints on the HOD parameters $M_{\mathrm{min}}, \, M_{\mathrm{min}, p}, \, M_{1}, \, M_{1, p}$.}. Fig.~\ref{fig:cls-best-fit} shows the theoretical predictions derived from maximum likelihood parameters together with our the measured power spectra. The corresponding minimum $\chi^{2}$ is $\chi^{2} = 82.0$. Computing the degrees of freedom as $\nu = N_{\mathrm{data}} - N_{\mathrm{param}} = 94 - 14 = 80$, we obtain $\chi^{2}_{\mathrm{red}} = \sfrac{\chi^{2}}{\nu} = 1.03$ ($p$-value $= 0.42$), which shows that the data are consistent with the best-fit theoretical model\footnote{We note that this estimate of the degrees of freedom is only valid for linear models and independent basis functions (see e.g. \cite{Andrae:2010}). However, we will use it throughout this work, as it allows us to obtain a rough estimate of the goodness of fit of the different models considered in our analysis.}.

@@

\begin{figure}
\begin{center}
\includegraphics[width=0.95\textwidth]{figures/Cls-data-vs-best-fit_mPk=HOD_fix=alpha-fc-sigmaM_HOD=zevol_fit=pz-shifts+prior=0p2-pz-widths+prior=0p2_fit=auto+cross_cosmo=const_cov=G+NG+SSC-LINBIAS_HOD-param=zfid_clfit=HOD-zevol_no-mag-vs-mag.pdf}
\caption{\todo{UPDATE WITH FULLY CONVERGED RESULTS} Measured auto- and cross power spectra considered in our analysis. The solid lines show the theoretical predictions obtained from the best-fit model parameters obtained in a joint fit to both auto- and cross power spectra not allowing for magnification while the dashed lines show the theoretical predictions obtained allowing for a free magnification amplitude $A_{\mu}$.}
\label{fig:cls-best-fit}
\end{center}
\end{figure}

@@

    From a joint fit to all auto- and cross power spectra we obtain $A_{\mu} = 0.99\substack{+0.37 \\ -0.36}$ and the constraint is shown in Fig.~\ref{fig:constraints-fit=mag-bias-ampl_fit=auto+cross-vs-fit=mag-bias-ampl_fit=auto}. This is consistent with our fiducial model for $s(z)$, which corresponds to $A_{\mu} = 1$ and constitutes a $\sim 2.8 \sigma$ detection of magnification from our HSC sample. The best-fit $\chi^{2}$ including magnification is $\chi^{2} = 83.8$\footnote{This is consistent with the case in which we set $A_{\mu}=1$, where we obtain consistent constraints with best-fit $\chi^{2} = 83.4$.}, which, computing the degrees of freedom as $\nu = N_{\mathrm{data}} - N_{\mathrm{param}} = 94 - 15 = 79$, leads to $\chi^{2}_{\mathrm{red}} = \sfrac{\chi^{2}}{\nu} = 1.06$ ($p$-value $= 0.33$). This corresponds to an improvement in $\chi^{2}$ compared to the fiducial analysis of $\Delta \chi^{2} = 95.8 - 83.8 = 12.0$, consistent with the $\sim 3 \sigma$ detection. Fig.~\ref{fig:cls-best-fit} shows the corresponding theoretical predictions derived from the maximum likelihood parameters alongside the observed data. The constraints on HOD parameters obtained when including magnification agree very well with our fiducial constraints, as can be seen from Fig.~\ref{fig:constraints-no-mag-bias-vs-fit=mag-bias-ampl}. In addition, we find the significance of the magnification detection to be largely insensitive to photo-$z$ systematics modeling choices, as it is not significantly affected by relaxing our photo-$z$ systematics model by fixing $z_{w, i} = 0$.

@@

    \begin{figure}
      \begin{center}
        \includegraphics[width=0.95\textwidth]{figures/contours-Mmin-M1_mPk=HOD_fix=alpha-fc-sigmaM_HOD=zevol_fit=pz-shifts+prior=0p2-pz-widths+prior=0p2_fit=auto+cross_cosmo=const-LINBIAS_HOD-param=zfid_clfit=HOD-zevol_no-mag-bias-vs-mag-bias-free.pdf}
        \caption{\todo{UPDATE WITH FULLY CONVERGED RESULTS} Comparison of our fiducial HOD constraints to those obtained when varying the magnification amplitude. The inner (outer) contour shows the $68 \%$ c.l. ($95 \%$ c.l.).}
        \label{fig:constraints-no-mag-bias-vs-fit=mag-bias-ampl}
      \end{center}
    \end{figure}

    \begin{figure}
      \begin{center}
        \includegraphics[width=0.3\textwidth]{figures/contours-A_mu_mPk=HOD_fix=alpha-fc-sigmaM_HOD=zevol_fit=pz-shifts+prior=0p2-pz-widths+prior=0p2_fit=auto+cross_cosmo=const-LINBIAS_HOD-param=zfid_clfit=HOD-zevol_mag-bias-free-fit=auto+cross-vs-mag-bias-free-fit=auto.pdf}
        \caption{\todo{UPDATE WITH FULLY CONVERGED RESULTS} Comparison of the constraints on magnification amplitude obtained from auto power spectra alone to those obtained from auto and cross power spectra. The solid line denotes the expectation value of the magnification amplitude determined from our data-driven fiducial model described in Sec.~\ref{ssec:results.magnification}.}
        \label{fig:constraints-fit=mag-bias-ampl_fit=auto+cross-vs-fit=mag-bias-ampl_fit=auto}
      \end{center}
    \end{figure}

@@

    To illustrate the cosmological constraining power of the data considered in our analysis, we allow for variations in the r.m.s. of linear matter fluctuations in spheres of comoving radius 8 $h^{-1}$ Mpc, $\sigma_{8}$, and the fractional cold matter density today, $\Omega_{c}$, in addition to our baseline set of parameters (c.f. Sec.~\ref{ssec:methods.constr}). All other cosmological parameters remain fixed to their fiducial Planck 2018 values. From a joint fit to all auto- and cross power spectra we obtain $\sigma_{8} = 0.837\substack{+0.193 \\ -0.211}$ and $\Omega_{c} = 0.228\substack{+0.022 \\ -0.023}$, consistent with our assumed fiducial cosmology. The corresponding constraints are shown in Fig.~\ref{fig:constraints-fit=Oc+s8}. As can be seen from Tab.~\ref{tab:constraints_robustness}, the derived constraints on HOD parameters agree well with our fiducial results, which shows that our HOD constraints are robust to variations in $\Omega_{c}$ and $\sigma_{8}$ as allowed by the data. This exploratory cosmological analysis further shows that the data are able to constrain $\Omega_{c}$ while $\sigma_{8}$ is largely unconstrained, showing that halo model is sufficiently flexible to allow effective marginalization over linear galaxy bias. Therefore, photometric current and future galaxy clustering analyses constitute a promising way to jointly constrain HOD and cosmological parameters, which we leave to future work.

>>>>

\abstract{We analyze the clustering of galaxies in the first public data release of the Hyper Suprime-Cam Subaru Strategic Program. Despite the relatively small footprints of the observed fields, the data are an excellent proxy for the very deep photometric datasets as will be acquired by the Large Synoptic Survey Telescope, and are therefore an ideal test bed for the analysis methods being implemented by the Dark Energy Science Collaboration. We select a magnitude limited sample with $i<24.5$ and analyze it in four tomographic redshift bins covering the range $0.15\lesssim z\lesssim1.5$. We carry out a Fourier-space analysis of the two-point clustering of this sample, including all auto- and cross-correlations between bins. We demonstrate the use of map-level deprojection methods to account for non-physical fluctuations in the galaxy number density caused by observational systematics. Through a halo occupation distribution analysis, we place constraints on the characteristic halo masses of this sample as a function of redshift, finding a good fit up to scales $k_{\rm max}=1\,{\rm Mpc}^{-1}$, including both auto- and cross-correlations. Our results show monotonically decreasing average halo masses, which can be interpreted in terms of the drop-out of red galaxies at high redshifts for a flux-limited sample, reproducing previous results. In terms of photometric redshift systematics, we show that additional care is needed in order to marginalize over uncertainties in the redshift distribution in galaxy clustering, even for samples of this small size, and that these uncertainties can be significantly constrained by including cross-bin correlations. We are able to make a $\sim3\sigma$ detection of the effects of lensing magnification in the HSC data. When freeing up cosmological parameters, we find constraints on the amplitude of density fluctuations $\sigma_8$ and the cold dark matter abundance $\Omega_c$ that agree well with measurements from Planck and low-redshift probes. Finally, we use our analysis pipeline to study the clustering of galaxies as a function of limiting flux, and provide a simple model for the linear galaxy bias for magnitude limited samples as a function of limiting magnitude and redshift.}

@@

The past two decades have seen a revolution in our understanding of the universe and its constituents. While observations of the cosmic microwave background play a pivotal role \cite{Planck:2018}, they are only indirectly sensitive to physics of the low redshift universe, in particular to the expansion history and growth, which are uniquely sensitive to properties of dark energy and modified gravity \cite{1903.12016}.

In studies of the low redshift universe, surveys of the large scale structure play an increasingly important role. Over a decade ago, the Dark Energy Task Force has divided the evolution of optical experiments into approximate stages \cite{0609591}. The current Stage III experiments, including spectroscopic surveys such as eBOSS \cite{1707.09322,1801.03038,1712.08064,1801.02689,1801.02891} and VIPERS \cite{1611.07048,1612.05645,1708.00026} and photometric surveys such as DES \cite{1708.01531,1708.01530,1708.01536}, Hyper Suprime-Cam \cite{1704.05858,2019PASJ...71...43H,1906.06041} and the Kilo-Degree Survey / VIKING-450 \cite{1902.11265,1812.06077,Joudaki:2019} are approaching completion and the field is preparing for the Stage IV surveys, such as DESI \cite{1611.00036,1907.10688} and LSST \cite{0912.0201,1809.01669} as well as major satellite missions like Euclid \cite{1606.00180} and WFIRST \cite{1904.01174}.

Together with transformational sensitivity increases in the Stage IV surveys, the challenges of understanding and controlling systematic errors are becoming considerably more difficult \cite{1808.07335}. For photometric surveys, the deep surveys sensitive to many more sources inevitably come with crowded fields where effects of blending affect a significant fraction of all sources in the field. The increased blending leads to new uncertainties in isolating sources, measuring their fluxes (required for photometric redshifts) and inferring their shear estimates. These effects in turn lead to subtle sample selection effects, which are amplified by the interaction of the point spread function with blending \cite{1708.01533,1905.01324,1907.10572}. But even with perfect measurements of fluxes, our ability to infer redshift distribution of sources is hampered by incompleteness in spectroscopic samples used to calibrate redshifts \cite{1903.09325}. Finally, at the depths of Stage IV photometric surveys, instrumental and observational effects, such as the scattered light from bright objects, errors in the galaxy-star separation, reddening by the milky way dust and others will impriment their own fluctuations into the galaxy density of objects. These fluctuations will, if not correctly accounted for, masquerade as intrinsic large-scale fluctuations in the cosmic density fields.  Therefore, Stage IV photometric experiments are likely to be limited by our ability to control systematics rather than the intrinsic statistical power of the survey.

Community is well aware of these challenges lying ahead. There are two main approaches that are being used in preparation for the arrival of data. On one hand, we are building very sophisticated realistic mock data sets, such as LSST DESC Data Challenge 2 \cite{1909.07340,1907.06530}. On the other hand, we are re-analyzing the existing precursor data to test-drive our methods and codes in realistic environment. In this paper we focus on the latter approach using Hyper Suprime-Cam Subaru Strategic Program (HSC-SSP) first data release  \cite{2018PASJ...70S...8A}. The main advantage of this dataset is that it is in many respects very similar to the LSST data, both in terms of depth and bands as well as primary data reduction and catalog generation. It covers essentially the same bands as LSST, with the exception of missing the UV $u$-band fluxes, which are unlikely to play a major role in cosmological data analysis. We will work with limiting magnitude for 5-$\sigma$ source detection of $i<24.5$, similar to expected magnitude cut to be used for LSST of $i<25.1$ after 1 year \cite{1809.01669}.  The primary data reduction codes used in HSC-SSP and planned for LSST are of the same lineage employing much of the same methods \cite{2018PASJ...70S...5B,1812.03248}. With the exception of a rather small sky area of around 90 square degrees used in this paper, the HSC-SSP data is a perfect example of how the LSST data will look like.

@@

The paper is structured as follows.  In Section \ref{sec:data}, we discuss the HSC-SSP data and how we process it in order to generate a set of galaxy over-density maps together with maps of potential systematics. In Section \ref{sec:methods}, we go through the methodology employed in our analysis and in Section \ref{sec:results}, we present results alongside numerous tests. We discuss and conclude in Section \ref{sec:discussion}.

@@

  HSC-SSP is a photometric galaxy survey that has been awarded 300 nights on the Subaru Telescope starting in 2014. DR1 includes data from 61.5 nights covering and comes with fields of three depths: Wide (108 square degrees to $i\sim 26.4$), Deep (26 square degrees to $i\sim26.5$) and UltraDeep (4 square degrees to $i\sim 27$). In this work we focus on the Wide field, which has been observed in five broadband filters ($grizy$) and is distributed among 7 fields between 5 and 20 square degrees each. The image quality is impressive at median seeing of around 0.6 arcsec, which is considerably better than other comparable surveys (e.g. DES with around 0.9 arc sec) and also likely better than median seeing that will be achieved by LSST. The data are processed with \texttt{hscPipe} \cite{2018PASJ...70S...5B} and  available to the community as a publicly available database.

@@

    \begin{table*}
      \caption{Summary of parameters varied in the MCMC with their respective priors. The posterior means and best-fits denote the values derived for our fiducial analysis. The uncertainties denote the $68 \%$ c.l..} \label{tab:params}
      \begin{center}
        \begin{tabular}{cccc}
          \hline\hline 
          Parameter & Prior & Posterior mean & Best-fit \\ \hline \Tstrut                             
          $M_{\mathrm{min}}$ & flat $\in [0., \,17.]$ & $11.88\substack{+0.22 \\ -0.23}$ & $12.02$ \\ 
          $M_{\mathrm{min}, p}$ & flat $\in [-10., \,10.]$ & $-0.491\substack{+2.09 \\ -2.02}$ & $-1.34$ \\
          $M_{0}$ & flat $\in [0., \,15.]$ & $5.72\substack{+3.95 \\ -3.97}$ & $6.60$ \\
          $M_{0, p}$ & flat $\in [-5., \,10.]$ & $2.45\substack{+5.09 \\ -5.02}$ & $-1.43$ \\
          $M_{1}$ & flat $\in [0., \,15.]$ & $13.08\substack{+0.27 \\ -0.28}$ & $13.27$ \\ 
          $M_{1, p}$ & flat $\in [-10., \,15.]$ & $0.933\substack{+2.67 \\ -2.56}$ & $-0.323$ \\
          $\Delta z_{0}$ & flat $\in [-0.2, \,0.2]$ & $0.000339\substack{+0.096 \\ -0.091}$ & $-0.0642$ \\
          $\Delta z_{1}$ & flat $\in [-0.2, \,0.2]$ & $-0.0159\substack{+0.096 \\ -0.086}$ & $-0.0865$ \\
          $\Delta z_{2}$ & flat $\in [-0.2, \,0.2]$ & $-0.00588\substack{+0.105 \\ -0.094}$ & $-0.0681$ \\
          $\Delta z_{3}$ & flat $\in [-0.2, \,0.2]$ & $0.0130\substack{+0.107 \\ -0.097}$ & $-0.0727$ \\
          $z_{w, 0}$ & flat $\in [-0.2, \,0.2]$ & $-0.0451\substack{+0.123 \\ -0.113}$ & $-0.101$ \\
          $z_{w, 1}$ & flat $\in [-0.2, \,0.2]$ & $0.00962\substack{+0.0780 \\ -0.0789}$ & $-0.0103$ \\
          $z_{w, 2}$ & flat $\in [-0.2, \,0.2]$ & $0.0346\substack{+0.0789 \\ -0.0778}$ & $0.0824$ \\
          $z_{w, 3}$ & flat $\in [-0.2, \,0.2]$ & $0.0543\substack{+0.112 \\ -0.126}$ & $0.182$ \\ \\
          $A_{\mu}$ & flat $\in [-5., \,5.]$ & $ - $ & $ - $ \\
          $\Omega_{c}$ & flat $\in [0.1, \,0.9]$ & $ - $ & $ - $ \\
          $\sigma_{8}$ & flat $\in [0.2, \,1.5]$ & $ - $ & $ - $ \\
          \hline\hline 
        \end{tabular}
      \end{center}
    \end{table*}

@@

Fig.~\ref{fig:constraints-fid-hod} shows the constraints on the HOD parameters $M_{\mathrm{min}}, \, M_{\mathrm{min}, p}, \, M_{1}, \, M_{1, p}$ for our fiducial model described in Sec.~\ref{ssec:methods.constr}. The constraints on all fitted parameters are shown in Fig.~\ref{fig:constraints-fid-full} and the corresponding best-fit values and means are shown alongside their $68 \%$ confidence limits in Tab.~\ref{tab:params}. As can be seen from Figures \ref{fig:constraints-fid-hod} and \ref{fig:constraints-fid-full}, the data allow us to constrain $M_{\mathrm{min}}(z)$ and $M_{1}(z)$ whereas $M_{0}(z)$ is unconstrained\footnote{In the following, we therefore only show constraints on the HOD parameters $M_{\mathrm{min}}, \, M_{\mathrm{min}, p}, \, M_{1}, \, M_{1, p}$.} \todo{explain upper bound}. Fig.~\ref{fig:cls-best-fit} shows the theoretical predictions derived from maximum likelihood parameters together with our the measured power spectra. The corresponding minimum $\chi^{2}$ is $\chi^{2} = 86.2$. Computing the degrees of freedom as $\nu = N_{\mathrm{data}} - N_{\mathrm{param}} = 94 - 14 = 80$, we obtain $\chi^{2}_{\mathrm{red}} = \sfrac{\chi^{2}}{\nu} = 1.08$ ($p$-value $= 0.30$), which shows that the data are consistent with the best-fit theoretical model\footnote{We note that this estimate of the degrees of freedom is only valid for linear models and independent basis functions (see e.g. \cite{Andrae:2010}). However, we will use it throughout this work, as it allows us to obtain a rough estimate of the goodness of fit of the different models considered in our analysis.}.

@@

\begin{figure}
\begin{center}
\includegraphics[width=0.95\textwidth]{figures/Cls-data-vs-best-fit_mPk=HOD_fix=alpha-fc-sigmaM_HOD=zevol_fit=pz-shifts+prior=0p2-pz-widths+prior=0p2_fit=auto+cross_cosmo=const_cov=G+NG+SSC-LINBIAS_HOD-param=zfid_clfit=HOD-zevol_no-mag-vs-mag.pdf}
\caption{Measured auto- and cross power spectra considered in our analysis. The solid lines show the theoretical predictions obtained from the best-fit model parameters obtained in a joint fit to both auto- and cross power spectra not allowing for magnification while the dashed lines show the theoretical predictions obtained allowing for a free magnification amplitude $A_{\mu}$.}
\label{fig:cls-best-fit}
\end{center}
\end{figure}

@@

    From a joint fit to all auto- and cross power spectra we obtain $A_{\mu} = 2.18 \pm 0.74$ and the constraint is shown in Fig.~\ref{fig:constraints-fit=mag-bias-ampl_fit=auto+cross-vs-fit=mag-bias-ampl_fit=auto}. This constitutes a $\sim 2.9 \sigma$ detection of magnification from our HSC sample. Furthermore, this constraint is consistent with our fiducial model for $s(z)$, which corresponds to $A_{\mu} = 1$, at the $\sim 1.6 \sigma$ level. This marginal inconsistency is probably due to uncertainties in the data-driven estimation of $s(z)$. The best-fit $\chi^{2}$ including magnification is $\chi^{2} = 69.0$\footnote{This is consistent with the case in which we set $A_{\mu}=1$, where we obtain consistent constraints with best-fit $\chi^{2} = 72.8$, as can be seen from Tab.~\ref{tab:constraints_robustness}.}, which, computing the degrees of freedom as $\nu = N_{\mathrm{data}} - N_{\mathrm{param}} = 94 - 15 = 79$, leads to $\chi^{2}_{\mathrm{red}} = \sfrac{\chi^{2}}{\nu} = 0.87$ ($p$-value $= 0.78$). This corresponds to an improvement in $\chi^{2}$ compared to the fiducial analysis of $\Delta \chi^{2} = 86.2 - 69.0 = 17.2$, consistent with the $\sim 3 \sigma$ detection. Fig.~\ref{fig:cls-best-fit} shows the corresponding theoretical predictions derived from the maximum likelihood parameters alongside the observed data. The constraints on HOD parameters obtained when including magnification agree very well with our fiducial constraints, as can be seen from Fig.~\ref{fig:constraints-no-mag-bias-vs-fit=mag-bias-ampl}. In addition, we find the significance of the magnification detection to be largely insensitive to photo-$z$ systematics modeling choices, as it is not significantly affected by relaxing our photo-$z$ systematics model by fixing $z_{w, i} = 0$.

@@

    \begin{figure}
      \begin{center}
        \includegraphics[width=0.95\textwidth]{figures/contours-Mmin-M1_mPk=HOD_fix=alpha-fc-sigmaM_HOD=zevol_fit=pz-shifts+prior=0p2-pz-widths+prior=0p2_fit=auto+cross_cosmo=const-LINBIAS_HOD-param=zfid_clfit=HOD-zevol_no-mag-bias-vs-mag-bias-free.pdf}
        \caption{Comparison of our fiducial HOD constraints to those obtained when varying the magnification amplitude. The inner (outer) contour shows the $68 \%$ c.l. ($95 \%$ c.l.).}
        \label{fig:constraints-no-mag-bias-vs-fit=mag-bias-ampl}
      \end{center}
    \end{figure}

    \begin{figure}
      \begin{center}
        \includegraphics[width=0.3\textwidth]{figures/contours-A_mu_mPk=HOD_fix=alpha-fc-sigmaM_HOD=zevol_fit=pz-shifts+prior=0p2-pz-widths+prior=0p2_fit=auto+cross_cosmo=const-LINBIAS_HOD-param=zfid_clfit=HOD-zevol_mag-bias-free-fit=auto+cross-vs-mag-bias-free-fit=auto.pdf}
        \caption{Comparison of the constraints on magnification amplitude obtained from auto power spectra alone to those obtained from auto and cross power spectra. The solid line denotes the expectation value of the magnification amplitude determined from our data-driven fiducial model described in Sec.~\ref{ssec:results.magnification}.}
        \label{fig:constraints-fit=mag-bias-ampl_fit=auto+cross-vs-fit=mag-bias-ampl_fit=auto}
      \end{center}
    \end{figure}

@@

    To illustrate the cosmological constraining power of the data considered in our analysis, we allow for variations in the r.m.s. of linear matter fluctuations in spheres of comoving radius 8 $h^{-1}$ Mpc, $\sigma_{8}$, and the fractional cold matter density today, $\Omega_{c}$, in addition to our baseline set of parameters (c.f. Sec.~\ref{ssec:methods.constr}). All other cosmological parameters remain fixed to their fiducial Planck 2018 values. From a joint fit to all auto- and cross power spectra we obtain $\sigma_{8} = 0.806\substack{+0.148 \\ -0.145}$ and $\Omega_{c} = 0.237 \pm 0.025$, \todo{UPDATE WITH FULLY CONVERGED RESULTS} consistent with our assumed fiducial cosmology. The corresponding constraints are shown in Fig.~\ref{fig:constraints-fit=Oc+s8}. As can be seen from Tab.~\ref{tab:constraints_robustness}, the derived constraints on HOD parameters agree well with our fiducial results, which shows that our HOD constraints are robust to variations in $\Omega_{c}$ and $\sigma_{8}$ as allowed by the data. This exploratory cosmological analysis further shows that the data are able to constrain $\Omega_{c}$ while $\sigma_{8}$ is largely unconstrained, showing that halo model is sufficiently flexible to allow effective marginalization over linear galaxy bias. Therefore, photometric current and future galaxy clustering analyses constitute a promising way to jointly constrain HOD and cosmological parameters, which we leave to future work.
